---
title: "Codes for raw data"
author: "Yunyi Sun"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### library
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(corrplot)
library(summarytools)
library(car)
library(stevemisc)
library(plyr)
### https://statisticsglobe.com/r-error-must-only-be-used-inside-dplyr-verbs 
## why R Error: `n()` must only be used inside dplyr verbs. (2 Examples)
# https://dplyr.tidyverse.org/reference/summarise.html
```


#### Data Import: Raw Data
```{r}

Combined_all3WAVES <- read.csv("../Combined__all3WAVES.csv",header=TRUE)
colnames(Combined_all3WAVES)

head(Combined_all3WAVES)

Combined_all3WAVES_text <- read.csv("../Combined_all3WAVES_text.csv",header=TRUE)

setdiff(colnames(Combined_all3WAVES_text),colnames(Combined_all3WAVES))

match(colnames(Combined_all3WAVES),colnames(Combined_all3WAVES_text))

names(Combined_all3WAVES)[colnames(Combined_all3WAVES) %nin% colnames(Combined_all3WAVES_text)]

```

#### Creating new Data File


#### Create subset of original datafile
```{r}
df1 <- Combined_all3WAVES %>% select(starts_with("Mental_Health_5_item_"),starts_with("Imposter_"),starts_with("Minority_Stress_"),Discipline,Race_ethnicity_ID,Gender,Sexual_orientation,U_others_sex_harass,Offended_sex_remarks,names(Combined_all3WAVES)[grepl("Institution", names(Combined_all3WAVES),fixed=TRUE)]) 
attach(df1)
head(df1)


temp1 <- names(Combined_all3WAVES)[grepl("Institution", names(Combined_all3WAVES),fixed=TRUE)]
temp2 <- names(Combined_all3WAVES_text)[grepl("Institution", names(Combined_all3WAVES_text),fixed=TRUE)]
for (i in temp1){
 print( unique(Combined_all3WAVES[i]))
  print(table(Combined_all3WAVES[i]))
  print(summary(Combined_all3WAVES[i]))
}


for (i in temp2){
 print( unique(Combined_all3WAVES_text[i]))
  print(table(Combined_all3WAVES_text[i]))
  print(summary(Combined_all3WAVES_text[i]))
}

Combined_all3WAVES_text %>% group_by(Institution_1) %>% dplyr::summarise(n=n())
Combined_all3WAVES_text %>% group_by(Institution_2) %>% dplyr::summarise(n=n())
sum(Combined_all3WAVES_text["Institution_2"][[1]]=="")
sum(is.na(Combined_all3WAVES_text["Institution_2"][[1]]))
Combined_all3WAVES_text[Combined_all3WAVES_text["Institution_2"]=="","Institution_2"] <- NA

# Combined_csv: in the column of "Institution_2", which indicates the specific institution of the respondents, in total we have respondents from 125 different institutions according to the answers to this question.   

# Subset data file: we have total 98 different institutions
df_Sexual %>% group_by(Institution_2) %>% dplyr::summarise(n=n()) 
df_Sexual %>% group_by(Institution_1) %>% dplyr::summarise(n=n())

table(df_Sexual$Institution_2)

```

#### Check the str of df1
```{r}
str(df1)
colnames(df1)

```

#### Recode the demographical variables 
```{r}

df1 <- df1 %>% 
  mutate(Discipline=recode(as.character(Discipline),
                      "c('1','2')='1';c('3','4','5')='2'"),Gender=recode(as.character(Gender),"c('3','4','5','6','7')='3'")) %>%
  mutate(across(c(Discipline,Gender,Sexual_orientation,
                  U_others_sex_harass,Offended_sex_remarks),~as.character(.))) %>%
  mutate_if(is.integer,as.numeric) 

#### check unique categories of recoded columns
unique(df1$Discipline)
unique(df1$Gender)
unique(df1$U_others_sex_harass)
unique(df1$Offended_sex_remarks)
```

#### Recode the Race since its subgroups are more complicated

```{r}
### Define a function to modify race column
Race_modify <-  function(x) x$Race_ethnicity_ID[i] <- ifelse("1" %in% str_split(x$Race_ethnicity_ID[i], ",", simplify = TRUE), "1",ifelse(x$Race_ethnicity_ID[i]!= "","2",NA)
 )

  for (i in 1:nrow(df1)) {
 df1$Race_ethnicity_ID[i]  <- Race_modify(df1)
  } 


#### check unique categories of recoded columns
unique(df1$Race_ethnicity_ID)
```

#### Recode the Mental health item 3 to 5 since they are coded reversely
```{r}

df1 <- df1 %>% mutate(across(paste("Mental_Health_5_item_",c(3,4,5),sep="")
                          ,~recode(as.character(.),"'1'='6';'2'='5';'3'='4';'4'='3';
                                      '5'='2';'6'='1'")))

#### check unique categories of recoded columns
head(unique(df1[paste("Mental_Health_5_item_",c(3,4,5),sep="")]))

```

#### Mutate all categorical variables to factor
```{r}
df1 <- df1 %>% mutate_if(is.character,as.factor) 

```


#### Label all categorical variables to factor

```{r}
# NOTE: BOTH dplyr and plyr has function rename, if you only call rename, R would automatically believe you are refering to the one from plyr, plyr::rename(d, c("old1"="new1", "old2"="new2")), which needed to be write like this
# df1 %>% plyr::rename(c(Discipline="Discipline_num",Gender="Gender_num",
#          Sexual_orientation="Sexual_orientation_num")) %>% colnames()

df1 <- df1 %>% dplyr::rename(Discipline_num=Discipline,Gender_num=Gender,
         Sexual_orientation_num=Sexual_orientation) %>% 
  mutate(
  Discipline=plyr::mapvalues(Discipline_num,c("1","2"),c("Stem","Non-Stem")),
  Gender=plyr::mapvalues(Gender_num,c("1","2","3"),c("Woman", "Man","Others")),
  Sexual_orientation=plyr::mapvalues(Sexual_orientation_num,c("1","2","3","4","5"),c("Hetero","Homo","Bi","NL","PNA")),
  Race=plyr::mapvalues(Race_ethnicity_ID,c("1","2"),c("Black","Non-Black")),
  Sex_harass=plyr::mapvalues(U_others_sex_harass,c("1","2","3"),c("Yes","No","Not_Sure")),
  Sex_remarks=plyr::mapvalues(Offended_sex_remarks,c("1","2","3"),c("Yes", "No","Not_Sure"))) %>% dplyr::rename(Race_num=Race_ethnicity_ID,Sex_harass_num=U_others_sex_harass,Sex_remarks_num=Offended_sex_remarks)


```


#### ADD per NA col to each Continuous variable
```{r}
idx_1<-sum(grepl("Mental_Health_5_item_", names(df1),fixed=TRUE))
idx_2<-sum(grepl("Imposter_", names(df1),fixed=TRUE))
idx_3<-sum(grepl("Minority_Stress_", names(df1),fixed=TRUE))

df1 <- df1 %>% mutate(perNA=rowSums(is.na(df1)),perNA_1 = rowSums(is.na(df1[grepl("Mental_Health_5_item_", names(df1),fixed=TRUE)])),perNA_2 = rowSums(is.na(df1[grepl("Imposter_", names(df1),fixed=TRUE)])),perNA_3 = rowSums(is.na(df1[grepl("Minority_Stress_", names(df1),fixed=TRUE)]))) %>% 
  relocate(perNA_1,.after=paste("Mental_Health_5_item_",idx_1,sep="")) %>% 
  relocate(perNA_2,.after=paste("Imposter_",idx_2,sep="")) %>% 
  relocate(perNA_3,.after=paste("Minority_Stress_",idx_3,sep=""))

attach(df1)

for (i in 1:3){
  print(table(eval(parse(text=paste('perNA_',i,sep="")))))
}
# or write get(paste('perNA_',i,sep=""))


```



#### Calculate the sum score for continuous varaible NOTE:na.rm = FALSE
```{r}

df1 <- df1 %>% transmute(!!!rlang::syms(setdiff(names(.), names(.)[c(which(grepl("Mental_Health_5_item_", names(df1),fixed=TRUE)),which(grepl("Imposter_",names(df1),fixed=TRUE)),which(grepl("Minority_Stress_",names(df1),fixed=TRUE)))])),Mental_Health_Score=select(.,starts_with("Mental_Health_5_item_")) %>% rowSums(),Imposter_Score=select(.,starts_with("Imposter_"))%>% rowSums(), Minority_Stress_Score=select(.,starts_with("Minority_Stress_")) %>% rowSums()) %>% 
  relocate(perNA_1,.after=paste("Mental_Health_","Score",sep="")) %>% 
  relocate(perNA_2,.after=paste("Imposter_","Score",sep="")) %>% 
  relocate(perNA_3,.after=paste("Minority_Stress_","Score",sep=""))


```


##### Creating a new categorical variable: Group based on conditions

```{r}


df1 <- df1 %>% mutate(Group = 
                 case_when(Sex_harass == "Yes" & Sex_remarks == "Yes" ~ "Both",
                           Sex_harass == "No" & Sex_remarks == "No" ~ "Neither",
                      xor (Sex_harass == "Yes", Sex_remarks =="Yes") ~ "Either" ,
                             TRUE ~ "Missing"))

df1$Group[df1$Group=="Missing"] <- NA
# df1$Group <- factor(df1$Group)
attach(df1)
# create a new perNA column
df_Sexual <- df1 %>% mutate(per_NA_Group = as.integer(is.na(df1["Group"]))) %>%
          filter(if_all(c(per_NA_Group,perNA_1,perNA_2,perNA_3), ~ (.x ==0)), apply(!is.na(df1[,which(names(df1) %in%                   c("Discipline","Gender","Race","Sexual_orientation","Group"))]),1,all)) %>% is.na() %>% rowSums() %>% sum()


# check if we still have NA
colSums(is.na(df_Sexual[,which(names(df_Sexual) %in%                   c("Discipline","Gender","Race","Sexual_orientation","Group"))]))



       # filter(if_all(all_of(c(Discipline, Race, Gender,Sexual_orientation,
       #                U_others_sex_harass,Offended_sex_remarks)), ~ !is.na(.x)))
```




#### Descriptive stats
```{r}
summary(df1)
dfSummary(df1[, !grepl("_num",colnames(df1))]) ##don't repeatedly calculate
########
# created a unique table for missing value col
df1 %>% group_by(perNA) %>% summarise(n=n())
# # make unique missing values as factor
# df1$perNA <- as.factor(df1$perNA)
# create a missing value dataset
dfmissing <- df1 %>% filter( perNA  > 0, apply(!is.na(df1[,which(names(df1) %in% c("Discipline","Gender","Race","Sexual_orientation","Group"))]),1,all)) 
head(dfmissing)


# create a cross-table to see the distribution of missing values across subcategories of one discrete variable

ctable(
  x = dfmissing$Discipline_num,
  y = dfmissing$perNA,
prop = "r")
ctable(
  x = dfmissing$Gender_num,
  y = dfmissing$perNA,
prop = "r")
ctable(
  x = dfmissing$Race_num,
  y = dfmissing$perNA,
prop = "r")
ctable(
  x = dfmissing$Sexual_orientation,
  y = dfmissing$perNA,
prop = "r")


```

#### Complement: You can write in a loop but not recommended
```{r}
temp <- names(Filter(is.factor,dfmissing))
for (i in 1:length(temp)){
  co_na <-temp[i]
  x1<-dfmissing[co_na][[1]]
  y1<-dfmissing["perNA"][[1]]
  t <-ctable(
  x = x1,
  y = y1,
prop = "r")
  print(t)
}
### you can try to write it within a loop but then you won't have name foace each table  

```
